Model Checking for Open Concurrent Systems

### 2018-10-11 2018-ModelCheckingPNets/checking.tex

% the name(s) of the author(s) follow(s) next
%
% NB: Chinese authors should write their first names(s) in front of
% their surnames. This ensures that the names appear correctly in
% the running heads and the author index.
%
\author{Tengfei?Li
\and Eric?Madelaine \and *
}
%
\authorrunning{Lecture Notes in Computer Science: Authors' Instructions}
% (feature abused for this document to repeat the title also on left hand pages)

@@

\begin{abstract}

The verification of concurrent systems, especially concurrent systems with data, have been researched for many years. However, the composition of concurrent systems have not been solved well. The major problem is ...

\keywords{ pM$\mu$\quad $pMG $\quad pLTS }
\end{abstract}

@@

 \subsubsection{Predicate}
 A transition in a pLTS is consisted of a tuple with labelling $guard$, $parameterised\; action$, $assignment$ between two states. Formally, a transition is defined as $s\xrightarrow{\langle\alpha, e_{b}, (x_{j}:=e_{j})^{j\in J}\rangle}s'$. While an open transition in an open automaton is defined with labelling $Pred$ and $Post$ between a composition of states in pLTSs and holes. In the open transition, $Pred$ is a predicate over the different variables of the terms, labels, and states. Specifically, a predicate consists of a set of parameterised actions in pLTSs and actions in holes. \NOTE{In a pNet, some ports from each component have an immediate relationship, and these ports are connected each other. Also, one of these ports is treated as a action to trigger a transition in the pLTS or in a hole. A predicate is a synchronization vector from the action in each pLTS and each hole.}

\subsection{opM$\mu$}
In order to check open automata, a new property language is present. The property language is called open parameterised modal $\mu$-calculus (opM$\mu$). We give the syntax in Tab.~\ref{opmu}:

\begin{table}
	%	\multirow{2}{*}{Algorithme}
	\setlength\tabcolsep{18pt}% default value is 6pt 
	\centering 
	\caption{Syntax of the opM$\mu$} \label{opmu}
	\begin{tabular}{>{\bfseries}c| c } 
		\hline 		
	Action formula of pLTS &     $\alpha ::= \{c(arg_{1}, ..., arg_{n})\mid arg_{i}=?x\mid\ !e\mid \tau\}$                  \\
	& $|\ \neg\alpha \ |\ \alpha_{1}\vee\alpha_{2}\ |\ \alpha_{1}\wedge\alpha_{2}$   \\ 	\hline

	Action formula of Hole &     $\beta ::= \{c(arg_{1}, ..., arg_{n})\mid arg_{j}=?x\mid\ !e\mid \tau\}$                  \\
	& $|\ \neg\beta \ |\ \beta_{1}\vee\beta_{2}\ |\ \beta_{1}\wedge\beta_{2}$   \\ 	\hline

	Action formula of pNet &     $\gamma ::= \{c\}\ |\ \neg\gamma \ |\ \gamma_{1}\vee\gamma_{2}\ |\ \gamma_{1}\wedge\gamma_{2}$   \\ 	\hline

	Predicate &     $pred ::= \alpha\wedge \beta\wedge \gamma$   \\ 	\hline

	 	&  $\varphi ::= p \mid X \mid \neg\varphi \mid \varphi_{1}\wedge\varphi_{2}\ |\ \varphi_{1}\vee\varphi_{2} $\\
	 State formula & $\ |\ [\alpha]\varphi \ |\ \langle\alpha\rangle\varphi \ |\ [\beta]\varphi \ |\ \langle\beta\rangle\varphi \ |\ [\gamma]\varphi \ |\ \langle\gamma\rangle\varphi $  \\ 
	& $\ |\ \exists x_{i}\varphi\ |\ \forall x_{i}\varphi\ |\ \mu X(e_{i}/x_{i})\varphi\ |\ \nu X(e_{i}/x_{i})\varphi$	\\ \hline 
	\end{tabular} 
\end{table}

\TODO{One thing is how to distinguish the action of pLTSs, Holes and pNet, i.e. whose action does $\langle resume\rangle$ stand for?}

An opM$\mu$ is defines as a property language that is used to describe the concurrent systems. $\alpha$ is an action formula from pLTS. $\beta$ is a Hole action formula. $\beta$ can be a $c(arg_i)$ with three kinds of parameter $?x, !e, \tau$,  or the composition of several $\beta$ formula with negation, disjunction and conjunction. $\gamma$ is an action formula from open transition, and it is defined as an action name or the negation of $gamma$, or the disjunction and conjunction of two $\gamma$ formula. A predicate, which is the disjunction of $\alpha$, $\beta$ and $\gamma$, labels an open transition. In the state formula, $[\beta]\varphi$, $\langle\beta\rangle\varphi$, $[\gamma]\varphi$ and $\langle\gamma\rangle\varphi$ are added into the list.

An example of opM$\mu$ is the formula for open automaton in Sec.~\ref{Sec2:usecase}.  The formula can be expressed as follows:
\begin{equation}\label{formula3}
	\nu X(\langle fail?M\rangle\mu Y(M/t)(\langle resume\rangle X \vee \langle tick\rangle Y(t-1/t)\vee \langle ask\wedge reset\rangle true))
\end{equation}
First, a $fail$ occurs with receiving a value $M$, and then a $tick$ happens with decreasing $t$ by one. Or the system returns true after an $ask$ and the subsequent $reset$ occur. Otherwise, the system restarts after a $resume$ happens.

@@

@@

\end{thebibliography}

>>>>

% the name(s) of the author(s) follow(s) next
%
% NB: Chinese authors should write their first names(s) in front of
% their surnames. This ensures that the names appear correctly in
% the running heads and the author index.
%
\author{Tengfei~Li
\and Eric~Madelaine \and *
}
%
\authorrunning{Lecture Notes in Computer Science: Authors' Instructions}
% (feature abused for this document to repeat the title also on left hand pages)

@@

\begin{abstract}
We present open pNets (open parameterised networks of processes) to describe a composition of concurrent systems. In order to verify the open pNets, in the paper, we present a new property language named opM$\mu$ (open parameterised modal $\mu$-calculus).

\keywords{ open pNets\quad opM$\mu$\quad pMG \quad  }
\end{abstract}

@@

 \subsubsection{Predicate}
 A transition in a pLTS is consisted as a tuple with labelling $guard$, $parameterised\; action$, $assignment$ between two states. Formally, a transition is defined as $s\xrightarrow{\langle\alpha, e_{b}, (x_{j}:=e_{j})^{j\in J}\rangle}s'$. While an open transition in an open automaton is defined with labelling $Pred$ and $Post$ between a composition of states in pLTSs and holes. In the open transition, $Pred$ is a predicate over the different variables of the terms, labels, and states. Specifically, a predicate consists of a set of parameterised actions in pLTSs and actions in holes. \NOTE{In a pNet, some ports from each component have an immediate relationship, and these ports are connected each other. Also, one of these ports is treated as a action to trigger a transition in the pLTS or in a hole. A predicate is a synchronisation vector from the action in each pLTS and each hole.}

\subsection{opM$\mu$}

 The property language is called open parameterised modal $\mu$-calculus (opM$\mu$). We give the syntax in Tab.~\ref{opmu}:

\begin{table}
	%	\multirow{2}{*}{Algorithme}
	\setlength\tabcolsep{18pt}% default value is 6pt 
	\centering 
	\caption{Syntax of the opM$\mu$} \label{opmu}
	\begin{tabular}{>{\bfseries}c| c } 
		\hline 
	Action expression & $e ::= \underline{c} \mid \{c(x)\mid x\in X\} \mid \{\beta_{j}\mid j\in J\}$\\ \hline
	Action formula & $\alpha ::= e \mid \neg\alpha\ \mid\ \alpha_{1}\vee\alpha_{2}\ \mid\ \alpha_{1}\wedge\alpha_{2}$   \\ 	\hline

	 State formula & $\varphi ::= \top \mid \bot \mid e_1 = e_2 \mid \neg\varphi \mid \varphi_{1}\wedge\varphi_{2}\ \mid \varphi_{1}\vee\varphi_{2} \mid [\alpha]\varphi$\\
	& $\mid \langle\alpha\rangle\varphi \mid \exists x_{i}\varphi \mid \forall x_{i}\varphi \mid \mu X(e_{i}/x_{i})\varphi \mid \nu X(e_{i}/x_{i})\varphi$	\\ \hline 
	\end{tabular} 
\end{table}

An opM$\mu$ is defined as a property language that is used to describe the concurrent systems. $\alpha$ is an action formula. $\tau$ is a silent action and $c$ is an action name. $c(x)$ is a parameterised action from holes. $\underline{c}$ is a global action from synchronisation vector, where $h_B$ is an action from a hole.

We propose a formula with opM$\mu$ based on the Open Automaton of the Failure Monitor architecture in Sec.~\ref{Sec2:usecase}. The formula can be expressed as follows:
\begin{equation}\label{formula3}
	\langle \underline{fail}\rangle \mu X(t)(\langle ask\rangle\langle reset\rangle \top \vee \langle tick\rangle X(t-1))
\end{equation}

If a $fail$ occurs with a hole action $h_B=fail(true)$, and then a $ask$ happens followed by a $reset$ action after $tick$ happens with decreasing $t$ by one. The formula can be employed to express a liveness property that an action $reset$ eventually happens.

Another example for the property of the open automaton of the Failure Monitor architecture is shown as follows:
\begin{equation}\label{formula4}
	\nu Y. \langle \underline{fail}\rangle \mu X(t).\langle ask\rangle\langle reset\rangle \top \vee (\langle \underline{resume}\rangle Y \vee \langle tick\rangle X(t-1))
\end{equation}

On all the path, a $\underline{fail}[h_B=fail(true)]\rangle$ happens, and then a $ask$ happens followed by a $reset$ action after $tick$ happens with decreasing $t$ by one, until $\langle \underline{resume}[h_B=resume(true)]\rangle Y$ holds after $M$ times $tick$ happens with decreasing $t$ by one.  The formula can be employed to express that an action $reset$ always eventually happens.

If we describe the Formula.~\ref{formula3} with a graphical pMG, it can be shown as Fig.~\ref{pmg3}:

\begin{figure}
\centering
	\begin{tikzpicture}[->, shorten >=2pt, >=stealth, node distance=1.2cm,
	noname/.style={ellipse, minimum width=1.5em, minimum height=1.5em, draw}]
	\node[noname] (1)                                            				 	{$t_{1}$};
	\node[noname] (2) [right=of 1]                               				 	{$t_{2}$};
	\node[noname] (3) [right=of 2]                              				 	{$t_{3}$};
	\node[noname] (4) [node distance=0.5cm and 2cm, above right=of 3]    {$t_{4}$};
	\node[noname] (5) [right=of 4]                                					{$t_{5}$};
	\node[noname] (6) [right=of 5]                                					{$t_{6}$};
	\node[noname] (7) [node distance=0.4cm and 1.8cm, below right=of 3]     {$t_{7}$};

	\path 	(1) edge                   	node[above] {} (2)
	(2) edge                   	node[above] {} (3)
	(3) edge 			node[above] {} (4)
	(4) edge    		node[below] {} (5)
	(5) edge 			node[above] {} (6)
	(3) edge    		node[below] {} (7)
	(7) edge[bend left]    node[below] {} (3);

	\end{tikzpicture}
	\caption{The associated pMG of Formula~\ref{formula3}}\label{pmg3}
\end{figure}

where:
\begin{itemize}
	\item $L_{N}(t_{1}) = (\{\}, \langle fail[h_B=fail(true)]\rangle, 1)$
	\item $L_{N}(t_{2}) = (\{t\}, t := M, 1)$
	\item $L_{N}(t_{3}) = (\{t\}, \vee, 1)$
	\item $L_{N}(t_{4}) = (\{t\}, \langle ask\rangle, 1)$
	\item $L_{N}(t_{5}) = (\{t\}, \langle reset\rangle, 1)$
	\item $L_{N}(t_{6}) = (\{\}, true, 1)$
	\item $L_{N}(t_{7}) = (\{t\}, \langle tick\rangle, 1)$
\end{itemize}

@@

An open automaton is unfolded by a pNet with some pLTSs and holes. An open transition a defined as a hole action, a predicate, a postcondition between a composition of states from different pLTSs and holes.

What we have to deal with in the algorithm is the $hole$ and $predicate$. \TODO{The detail}

\subsubsection{Processing Holes}
We change the algorithm to make it suitable for checking open pNets. From the syntax of opM$\mu$, we define the global action including holes in the action as $\underline{c}(h_B = c(x))$. We change the modality to $\langle c\rangle$, $[c]$, $\langle \underline{c}(h_B = c(x))\rangle$, $[\underline{c}(h_B = c(x))]$.
It was shown as follows:

$\langle \underline{c}(h_B = c(x))\rangle\Rightarrow$
	$ checkAnd(\{ (p_{i}[v/y], (n',\rho\{x \rightarrow v\}))\,|\,  n \rightarrow n', p \xrightarrow{\langle \underline{c}(h_B = c(x))\rangle } p_{i},v\in Val\})$

{$[\underline{c}(h_B = c(x))]\Rightarrow$}
	 $ checkOr(\{ (p_{i}[v/y], (n', \rho\{x \rightarrow v\}))\,|\,  n \rightarrow n', p \xrightarrow{[\underline{c}(h_B = c(x))]} p_{i},v\in Val\})$

@@

	\bibitem{henrio2016pnet}Henrio, Ludovic, Eric Madelaine, and Min Zhang. "A theory for the composition of concurrent processes." International Conference on Formal Techniques for Distributed Objects, Components, and Systems. Springer, Cham, 2016.
\end{thebibliography}
